\documentclass{article}

\usepackage[top=4cm, bottom=5cm]{geometry}

\usepackage{amsmath,amssymb,amsfonts,amsthm,fontawesome}

\usepackage[english]{babel}
\usepackage{pagecolor,lipsum,hyperref}
\hypersetup{ colorlinks=true,
			 linkcolor=blue!70!red,
			 urlcolor=blue!70!red}

\usepackage{enumitem}
\setlist[itemize]{noitemsep,topsep=2pt}
\pagecolor{blue!15}
\usepackage{lmodern}
\def\titlefotn{\textsf}
% \usepackage{fontspec}

% \setmainfont{DejaVuSansMono-Bold}
% \newfontfamily{\titlefotn}{Futura-Condensed-Bold}
\pagestyle{empty}
\begin{document}
\begin{center}
\underline{\huge\titlefotn{Categorical tools}}\\[3em]
\Large \titlefotn {M a s a r y k \hspace{5pt} U n i v e r s i t y}\\
\titlefotn{S p r i n g \hspace{5pt} 2 0 1 7}
\end{center}

\vspace*{2em}
Dear students,

we are a postdoc (Fosco) and a PhD student (Ivan) here in MU; we are planning to start a series of lectures on two topics we love:
\begin{itemize}
	\item \textbf{Accessible and presentable} categories; 
	\item \textbf{Homotopical algebra} and abstract homotopy theory.
\end{itemize}

If you wanna be fashionable, and know some cutting-edge Mathematics, drop us a line!

You should be familiar with some rudiments of category theory, but we are trying to design the course in order to keep its prerequisites to a reasonable minimum. Of course, everyone willing to collaborate is welcome.

With the exception of the first lecture, we meet every Tuesday at 4pm in room XYZ.

Other useful informations:
\begin{itemize}
\item[\faCalendarCheckO \; :] Here's a tentative schedule of the first meetings:
\begin{itemize}
\item \textbf{6th April}: 
\begin{enumerate}
\item Ivan: Generalities on locally presentable categories. 
\item Fosco: Preliminaries and generalities on model categories.
\end{enumerate}
\item \textbf{11th April}:
\begin{enumerate}
\item Ivan: Representation theorem. 
\item Fosco: A model structure on simplicial sets and spaces: a Quillen equivalence.
\end{enumerate} 
\item \textbf{18th April}: 
\begin{enumerate}
\item Ivan: Properties of Locally Presentable Categories. 
\item Fosco: Homotopy Limits.
\end{enumerate}
\end{itemize}
\item[\faEnvelopeO \; :] \href{mailto:fosco.loregian@gmail.com}{\tt fosco.loregian@gmail.com} and/or \href{mailto:diliberti@math.muni.cz}{\tt diliberti@math.muni.cz}.
\item[\faGithub \; :] There's a github repo where we host material and relevant infos: 
\begin{center}\href{https://github.com/tetrapharmakon/progetti-MUNI}{\tt https://github.com/tetrapharmakon/progetti-MUNI}
\end{center}
\item
\end{itemize}
\begin{flushright}
Fosco and Ivan,\\ 
Brno, \today
\end{flushright}
\end{document}